---
title: "Untitled"
author: "Shelly Trigg"
date: "3/27/2020"
output: html_document
---

# Part 1:  Filter methylated regions for regions that have coverage in 75% of samples per group


manually removed the "#" in front the chromosome column name before reading in files

read in ambient DMR table
```{r}
amb_DMR <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20200327/amb_AllTimes_DMR250bp_cov5x_rms_results_collapsed.tsv", header = TRUE,sep = "\t",stringsAsFactors = FALSE)
#loop through table and keep only lines where up to one sample contains NA for % methylation

df <- data.frame() #create empty data frame to bind filtered rows into
for(i in (1:nrow(amb_DMR))){
  day0 <- amb_DMR[i,7:10] #define columns from the category Day 0
  day10 <- amb_DMR[i,11:14] #define columns from the category Day 10
  day135 <- amb_DMR[i,15:18] #define columns from the category Day 135
  day145 <- amb_DMR[i,19:22] #define columns from the category Day 145
  if(length(which(is.na(day0))) < 2 & length(which(is.na(day10))) < 2 & length(which(is.na(day135))) < 2 & length(which(is.na(day145))) < 2){
  df <- rbind(df,amb_DMR[i,]) #conditional statement: if less than 2 sameples/category have NA for % methylation bind the whole row to the new dataframe
  }
}
#write the output
write.table(df,"amb_AllTimes_DMR250bp_cov5x_rms_results_filtered.tsv", sep = "\t", row.names= FALSE, quote = FALSE)

#do the same for day 10

day10_DMR <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20200327/day10_AllpH_DMR250bp_cov5x_rms_results_collapsed.tsv", header = TRUE,sep = "\t",stringsAsFactors = FALSE)
#loop through table and keep only lines where up to one sample contains NA for % methylation

df <- data.frame() #create empty data frame to bind filtered rows into
for(i in (1:nrow(day10_DMR))){
  LowpH <- day10_DMR[i,c(7,8,15,16)] #define columns from the category low pH
  SuperLowpH <- day10_DMR[i,c(9,10,17,18)] #define columns from the category super low pH
  amb <- day10_DMR[i,c(11:14)] #define columns from the category ambient
  if(length(which(is.na(LowpH))) < 2 & length(which(is.na(SuperLowpH))) < 2 & length(which(is.na(amb))) < 2){
    df <- rbind(df,day10_DMR[i,]) #conditional statement: if less than 2 sameples/category have NA for % methylation bind the whole row to the new dataframe
  }
}
#write the output
write.table(df,"day10_AllpH_DMR250bp_cov5x_rms_results_filtered.tsv", sep = "\t", row.names= FALSE, quote = FALSE)

#do the same for day 135

day135_DMR <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20200327/day135_AllpH_DMR250bp_cov5x_rms_results_collapsed.tsv", header = TRUE,sep = "\t",stringsAsFactors = FALSE)
#loop through table and keep only lines where up to one sample contains NA for % methylation

df <- data.frame() #create empty data frame to bind filtered rows into
for(i in (1:nrow(day135_DMR))){
  amb <- day135_DMR[i,7:10] #define columns from the category low pH
  LowpH <- day135_DMR[i,11:14] #define columns from the category super low pH
  SuperLowpH <- day135_DMR[i,15:18] #define columns from the category ambient
  if(length(which(is.na(LowpH))) < 2 & length(which(is.na(SuperLowpH))) < 2 & length(which(is.na(amb))) < 2){
    df <- rbind(df,day135_DMR[i,]) #conditional statement: if less than 2 sameples/category have NA for % methylation bind the whole row to the new dataframe
  }
}
#write the output
write.table(df,"day135_AllpH_DMR250bp_cov5x_rms_results_filtered.tsv", sep = "\t", row.names= FALSE, quote = FALSE)

#do the same for day 145

day145_DMR <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20200327/day145_AllpH_DMR250bp_cov5x_rms_results_collapsed.tsv", header = TRUE,sep = "\t",stringsAsFactors = FALSE)
#loop through table and keep only lines where up to one sample contains NA for % methylation

df <- data.frame() #create empty data frame to bind filtered rows into
for(i in (1:nrow(day145_DMR))){
  AmbAmb <- day145_DMR[i,9:12] #define columns from the category low pH
  LowAmb <- day145_DMR[i,c(7,8,15,16)] #define columns from the category super low pH
  SuperLowAmb <- day145_DMR[i,c(13,14,17,18)] #define columns from the category ambient
  AmbLow <- day145_DMR[i,c(19,20,27,28)] #define columns from the category ambient
  LowLow <- day145_DMR[i,c(21,22,29,30)] #define columns from the category ambient
  SuperLowLow <- day145_DMR[i,23:26] #define columns from the category ambient
  if(length(which(is.na(AmbAmb))) < 2 & length(which(is.na(LowAmb))) < 2 & length(which(is.na(SuperLowAmb))) < 2 & length(which(is.na(AmbLow))) < 2 & length(which(is.na(LowLow))) < 2 & length(which(is.na(SuperLowLow))) < 2){
    df <- rbind(df,day145_DMR[i,]) #conditional statement: if less than 2 sameples/category have NA for % methylation bind the whole row to the new dataframe
  }
}
#write the output
write.table(df,"day145_AllpH_DMR250bp_cov5x_rms_results_filtered.tsv", sep = "\t", row.names= FALSE, quote = FALSE)
```

# Part 2: Run group statistics on regions to find regions that are significantly different among groups

load libraries
```{r}
library(gplots)
library(ggplot2)
library(dplyr)
library(broom)
library(RColorBrewer)
```

read in data
```{r}

ambDMRs <- read.table("amb_AllTimes_DMR250bp_cov5x_rms_results_filtered.tsv", header = TRUE, sep = "\t")

d10DMRs <- read.table("day10_AllpH_DMR250bp_cov5x_rms_results_filtered.tsv", header = TRUE, sep = "\t")


d135DMRs <- read.table("day135_AllpH_DMR250bp_cov5x_rms_results_filtered.tsv", header = TRUE, sep = "\t")


d145DMRs <- read.table("day145_AllpH_DMR250bp_cov5x_rms_results_filtered.tsv", header = TRUE, sep = "\t")

meta_data <- read.csv("Sample.Info.csv", stringsAsFactors = FALSE)

```

Make a unique ID column 
```{r}
#for all ambient sample comparison
d10DMRs$ID <- paste(d10DMRs$chr,":",d10DMRs$start,"-",d10DMRs$end, sep = "")
d10DMRs$ID <- gsub("__.*__.*:",":",d10DMRs$ID)

d135DMRs$ID <- paste(d135DMRs$chr,":",d135DMRs$start,"-",d135DMRs$end, sep = "")
d135DMRs$ID <- gsub("__.*__.*:",":",d135DMRs$ID)

ambDMRs$ID <- paste(ambDMRs$chr,":",ambDMRs$start,"-",ambDMRs$end, sep = "")
ambDMRs$ID <- gsub("__.*__.*:",":",ambDMRs$ID)

d145DMRs$ID <- paste(d145DMRs$chr,":",d145DMRs$start,"-",d145DMRs$end, sep = "")
d145DMRs$ID <- gsub("__.*__.*:",":",d145DMRs$ID)

```

reformat data for calculating group effect
```{r}
#reformat all to long format
d10DMRs_STACKED <- tidyr::gather(d10DMRs[,7:ncol(d10DMRs)], "Sample.ID", "perc.meth",1:12)

d135DMRs_STACKED <- tidyr::gather(d135DMRs[,7:ncol(d135DMRs)], "Sample.ID", "perc.meth",1:12)

d145DMRs_STACKED <- tidyr::gather(d145DMRs[,7:ncol(d145DMRs)], "Sample.ID", "perc.meth",1:24)

ambDMRs_STACKED <- tidyr::gather(ambDMRs[,7:ncol(ambDMRs)], "Sample.ID", "perc.meth",1:16)


#simplify sample ID column
d10DMRs_STACKED$Sample.ID <- gsub("methylation_level_","", d10DMRs_STACKED$Sample.ID)
d10DMRs_STACKED$Sample.ID <- gsub("\\.","_", d10DMRs_STACKED$Sample.ID)

d135DMRs_STACKED$Sample.ID <- gsub("methylation_level_","", d135DMRs_STACKED$Sample.ID)
d135DMRs_STACKED$Sample.ID <- gsub("\\.","_", d135DMRs_STACKED$Sample.ID)

d145DMRs_STACKED$Sample.ID <- gsub("methylation_level_","", d145DMRs_STACKED$Sample.ID)
d145DMRs_STACKED$Sample.ID <- gsub("\\.","_", d145DMRs_STACKED$Sample.ID)

ambDMRs_STACKED$Sample.ID <- gsub("methylation_level_","", ambDMRs_STACKED$Sample.ID)
ambDMRs_STACKED$Sample.ID <- gsub("\\.","_", ambDMRs_STACKED$Sample.ID)

#merge with meta data
d10DMRs_STACKED <- merge(d10DMRs_STACKED, meta_data, by = "Sample.ID")

d135DMRs_STACKED <- merge(d135DMRs_STACKED, meta_data, by = "Sample.ID")

d145DMRs_STACKED <- merge(d145DMRs_STACKED, meta_data, by = "Sample.ID")
#create column for initial and secondary treatment
d145DMRs_STACKED$Treatment <- paste(d145DMRs_STACKED$Initial.Treatment,d145DMRs_STACKED$Secondary.Treatment, sep ="_")

ambDMRs_STACKED <- merge(ambDMRs_STACKED, meta_data, by = "Sample.ID")

```

plot % meth dist.
```{r}
jpeg("d10DMR_percmeth_hist.jpg", width = 800, height = 800)
ggplot(d10DMRs_STACKED) + geom_histogram(aes(perc.meth, group = Initial.Treatment, color = Initial.Treatment,fill = Initial.Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()

jpeg("d135DMR_percmeth_hist.jpg", width = 800, height = 800)
ggplot(d135DMRs_STACKED) + geom_histogram(aes(perc.meth, group = Initial.Treatment, color = Initial.Treatment,fill = Initial.Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()

jpeg("d145DMR_percmeth_hist.jpg", width = 800, height = 800)
ggplot(d145DMRs_STACKED) + geom_histogram(aes(perc.meth, group = Treatment, color = Treatment,fill = Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()

jpeg("ambDMR_percmeth_hist.jpg", width = 800, height = 800)
ggplot(ambDMRs_STACKED) + geom_histogram(aes(perc.meth, group = TimePoint, color = TimePoint,fill = TimePoint), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()
```

arc sin sqrt transform the data
```{r}
#arcsin sqrt transformation function
asinTransform <- function(p) { asin(sqrt(p))}

#arcsin transform data 
#day 10
d10DMRs_STACKED_asin <- d10DMRs_STACKED
d10DMRs_STACKED_asin$perc.meth <- asinTransform(d10DMRs_STACKED_asin$perc.meth)
#day 135
d135DMRs_STACKED_asin <- d135DMRs_STACKED
d135DMRs_STACKED_asin$perc.meth <- asinTransform(d135DMRs_STACKED_asin$perc.meth)
#day145
d145DMRs_STACKED_asin <- d145DMRs_STACKED
d145DMRs_STACKED_asin$perc.meth <- asinTransform(d145DMRs_STACKED_asin$perc.meth)
#amb samples
ambDMRs_STACKED_asin <- ambDMRs_STACKED
ambDMRs_STACKED_asin$perc.meth <- asinTransform(ambDMRs_STACKED_asin$perc.meth)


```

plot distribution of TRANSFORMED % methylation in all DMRs in all samples
```{r}
jpeg("d10DMR_Tpercmeth_hist.jpg", width = 800, height = 800)
ggplot(d10DMRs_STACKED_asin) + geom_histogram(aes(perc.meth, group = Initial.Treatment, color = Initial.Treatment,fill = Initial.Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()

jpeg("d135DMR_Tpercmeth_hist.jpg", width = 800, height = 800)
ggplot(d135DMRs_STACKED_asin) + geom_histogram(aes(perc.meth, group = Initial.Treatment, color = Initial.Treatment,fill = Initial.Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()

jpeg("d145DMR_Tpercmeth_hist.jpg", width = 800, height = 800)
ggplot(d145DMRs_STACKED_asin) + geom_histogram(aes(perc.meth, group = Treatment, color = Treatment,fill = Treatment), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()

jpeg("ambDMR_Tpercmeth_hist.jpg", width = 800, height = 800)
ggplot(ambDMRs_STACKED_asin) + geom_histogram(aes(perc.meth, group = TimePoint, color = TimePoint,fill = TimePoint), bins = 10, position = "identity", alpha = 0.5) + theme_bw()
dev.off()
```

## Run anova on TRANSFORMED data to assess group differences for each DMR

**Hypotehsis: there is no effect from pH at day10**
```{r}
d10DMR_1way_aov_pH <- d10DMRs_STACKED_asin %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~Initial.Treatment, data =  . ))
#summarize ANOVA data
d10DMR_1way_aov_pH_modelsumm <- glance(d10DMR_1way_aov_pH, meth_aov_models)
#write ANOVA output to file
write.csv(d10DMR_1way_aov_pH_modelsumm, "d10DMR_1way_aov_pH_modelsumm.csv", row.names = FALSE, quote = FALSE)

#run Tukey posthoc
d10DMR_1way_tuk <- d10DMRs_STACKED_asin %>% group_by(ID) %>%
do(meth_tuk_models = TukeyHSD(aov(perc.meth~Initial.Treatment, data =  . )))
#summarize Tukey results
d10DMR_1way_tuk_modelsumm <- tidy(d10DMR_1way_tuk, meth_tuk_models)
#reformat table to spread
d10DMR_1way_tuk_modelsumm <- tidyr::spread(d10DMR_1way_tuk_modelsumm[,c(1,2,3,7)],"comparison","adj.p.value")


```

**Hypotehsis: there is no effect from pH at day135**
```{r}
d135DMR_1way_aov_pH <- d135DMRs_STACKED_asin %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~Initial.Treatment, data =  . ))
#summarize ANOVA data
d135DMR_1way_aov_pH_modelsumm <- glance(d135DMR_1way_aov_pH, meth_aov_models)
#write ANOVA output to file
write.csv(d135DMR_1way_aov_pH_modelsumm, "d135DMR_1way_aov_pH_modelsumm.csv", row.names = FALSE, quote = FALSE)
```

**Hypotehsis: there is no effect from pH at day145**
```{r}
d145DMR_1way_aov_pH <- d145DMRs_STACKED_asin %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~Treatment, data =  . ))
#summarize ANOVA data
d145DMR_1way_aov_pH_modelsumm <- glance(d145DMR_1way_aov_pH, meth_aov_models)
#write ANOVA output to file
write.csv(d145DMR_1way_aov_pH_modelsumm, "d145DMR_1way_aov_pH_modelsumm.csv", row.names = FALSE, quote = FALSE)
```


**Hypotehsis: there is no effect from time in ambient samples**
```{r}
ambDMR_1way_aov_pH <- ambDMRs_STACKED_asin %>% group_by(ID) %>%
do(meth_aov_models = aov(perc.meth~TimePoint, data =  . ))
#summarize ANOVA data
ambDMR_1way_aov_pH_modelsumm <- glance(ambDMR_1way_aov_pH, meth_aov_models)
#write ANOVA output to file
write.csv(ambDMR_1way_aov_pH_modelsumm, "ambDMR_1way_aov_pH_modelsumm", row.names = FALSE, quote = FALSE)
```



plot significant regions
```{r}

### DAY 10 ###

jpeg("d10DMR_Taov0.05pHPercMeth.jpg", width = 17, height = 12, units = "in", res = 300)
p <- ggplot(data = d10DMRs_STACKED[which(d10DMRs_STACKED$ID %in% pull(d10DMR_1way_aov_pH_modelsumm[which(d10DMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),],aes(x = Initial.Treatment,y = perc.meth, color = Initial.Treatment)) + facet_wrap(~ID, scale = "free") + theme_bw() + theme(axis.text.x = element_text(size = 7,angle = 45, hjust =1),axis.title=element_text(size=12,face="bold")) + scale_color_manual(values = c("cyan3", "gold2", "magenta3"))
p + geom_jitter(width = 0.3) +  ggtitle("DMRs that show an overall pH model effect significant at ANOVA p.value < 0.05")
dev.off()

### DAY 135 ###

jpeg("d135DMR_Taov0.05pHPercMeth.jpg", width = 17, height = 12, units = "in", res = 300)
p <- ggplot(data = d135DMRs_STACKED[which(d135DMRs_STACKED$ID %in% pull(d135DMR_1way_aov_pH_modelsumm[which(d135DMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),],aes(x = Initial.Treatment,y = perc.meth, color = Initial.Treatment)) + facet_wrap(~ID, scale = "free") + theme_bw() + theme(axis.text.x = element_text(size = 7,angle = 45, hjust =1),axis.title=element_text(size=12,face="bold")) + scale_color_manual(values = c("cyan3", "gold2", "magenta3"))
p + geom_jitter(width = 0.3) +  ggtitle("Day 135 DMRs that show an overall pH model effect significant at ANOVA p.value < 0.05")
dev.off()

### DAY 145 ###
jpeg("d145DMR_Taov0.05pHPercMeth.jpg", width = 17, height = 12, units = "in", res = 300)
p <- ggplot(data = d145DMRs_STACKED[which(d145DMRs_STACKED$ID %in% pull(d145DMR_1way_aov_pH_modelsumm[which(d145DMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),],aes(x =Treatment,y = perc.meth, color = Treatment)) + facet_wrap(~ID, scale = "free") + theme_bw() + theme(axis.text.x = element_text(size = 7,angle = 45, hjust =1),axis.title=element_text(size=12,face="bold")) + scale_color_manual(values = c("cyan3","green3","gold2","greenyellow", "magenta3","lightslateblue"))
p + geom_jitter(width = 0.3) +  ggtitle("Day 145 DMRs that show an overall pH model effect significant at ANOVA p.value < 0.05")
dev.off()

### Amb samples ###
jpeg("ambDMR_Taov0.05pHPercMeth.jpg", width = 17, height = 12, units = "in", res = 300)
p <- ggplot(data = ambDMRs_STACKED[which(ambDMRs_STACKED$ID %in% pull(ambDMR_1way_aov_pH_modelsumm[which(ambDMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),],aes(x =TimePoint,y = perc.meth, color = TimePoint)) + facet_wrap(~ID, scale = "free") + theme_bw() + theme(axis.text.x = element_text(size = 7,angle = 45, hjust =1),axis.title=element_text(size=12,face="bold"))
p + geom_jitter(width = 0.3) +  ggtitle("Ambient x Time DMRs that show an overall pH model effect significant at ANOVA p.value < 0.05")
dev.off()
```

plot heatmap of 1way aov for pH p.val 0.05 sig DMRs
```{r}
#create matrix for d10 samples
aov_0.05pH_d10DMR_m <- as.matrix(d10DMRs[,7:18])
rownames(aov_0.05pH_d10DMR_m) <- d10DMRs$ID

aov_0.05pH_d10DMR_m <- aov_0.05pH_d10DMR_m[which(rownames(aov_0.05pH_d10DMR_m) %in% pull(d10DMR_1way_aov_pH_modelsumm[which(d10DMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),]
#remove extra text from column names
colnames(aov_0.05pH_d10DMR_m) <- gsub("methylation_level_","",colnames(aov_0.05pH_d10DMR_m))


#reorder columns by pH treatment (amb, low, superlow)
aov_0.05pH_d10DMR_m <- aov_0.05pH_d10DMR_m[,c(5,6,7,8,1,2,9,10,3,4,11,12)]
#define color bar corresponding to treatment
ColSideColors <- cbind(pH = c(rep("cyan3",4),rep("gold2",4),rep("magenta3",4)))

jpeg("DMR_Taov0.05_d10_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_d10DMR_m,margins = c(10,20), cexRow = 1.2, cexCol = 1, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)), ColSideColors = ColSideColors,na.color = "black", density.info = "none", trace = "none", scale = "row", sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_d10DMR_m),rowsep=1:nrow(aov_0.05pH_d10DMR_m))
dev.off()

###Visualize group means
#calculate group means
MeanD10pHamb <- rowMeans(aov_0.05pH_d10DMR_m[,grep("119|120|135|136", colnames(aov_0.05pH_d10DMR_m))], na.rm = TRUE)

MeanD10pHlow <- rowMeans(aov_0.05pH_d10DMR_m[,grep("103|104|127|128", colnames(aov_0.05pH_d10DMR_m))], na.rm = TRUE)

MeanD10pHslow <- rowMeans(aov_0.05pH_d10DMR_m[,grep("111|113|143|145", colnames(aov_0.05pH_d10DMR_m))], na.rm = TRUE)

#bind all group means together
aov_0.05pH_d10DMR_mean_m <- as.matrix(data.frame(cbind(MeanD10pHamb,MeanD10pHlow, MeanD10pHslow)))
#define color bar for group means
ColSideColors <- cbind(pH = c(rep("cyan3",1),rep("gold2",1),rep("magenta3",1)))

#plot group means heatmap
jpeg("DMR_Taov0.05_d10mean_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_d10DMR_mean_m,margins = c(10,20), cexRow = 1.2, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)), distfun = function(x) as.dist(1-cor(t(x))), hclustfun = function(x) hclust(x, method="average"),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_d10DMR_mean_m),rowsep=1:nrow(aov_0.05pH_d10DMR_mean_m))
dev.off()



### DAY 135 ###
#create matrix for d135 samples
aov_0.05pH_d135DMR_m <- as.matrix(d135DMRs[,7:18])
rownames(aov_0.05pH_d135DMR_m) <- d135DMRs$ID

aov_0.05pH_d135DMR_m <- aov_0.05pH_d135DMR_m[which(rownames(aov_0.05pH_d135DMR_m) %in% pull(d135DMR_1way_aov_pH_modelsumm[which(d135DMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),]
#remove extra text from column names
colnames(aov_0.05pH_d135DMR_m) <- gsub("methylation_level_","",colnames(aov_0.05pH_d135DMR_m))


#columns are already ordered by pH treatment (amb, low, superlow)
#define color bar corresponding to treatment
ColSideColors <- cbind(pH = c(rep("cyan3",4),rep("gold2",4),rep("magenta3",4)))

jpeg("DMR_Taov0.05_d135_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_d135DMR_m,margins = c(10,20), cexRow = 1.2, cexCol = 1, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)), ColSideColors = ColSideColors,na.color = "black", density.info = "none", trace = "none", scale = "row", sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_d135DMR_m),rowsep=1:nrow(aov_0.05pH_d135DMR_m))
dev.off()

###Visualize group means
#calculate group means
Meand135pHamb <- rowMeans(aov_0.05pH_d135DMR_m[,grep("151|152|153|154", colnames(aov_0.05pH_d135DMR_m))], na.rm = TRUE)

Meand135pHlow <- rowMeans(aov_0.05pH_d135DMR_m[,grep("159|160|161|162", colnames(aov_0.05pH_d135DMR_m))], na.rm = TRUE)

Meand135pHslow <- rowMeans(aov_0.05pH_d135DMR_m[,grep("167|168|169|170", colnames(aov_0.05pH_d135DMR_m))], na.rm = TRUE)

#bind all group means together
aov_0.05pH_d135DMR_mean_m <- as.matrix(data.frame(cbind(Meand135pHamb,Meand135pHlow, Meand135pHslow)))
#define color bar for group means
ColSideColors <- cbind(pH = c(rep("cyan3",1),rep("gold2",1),rep("magenta3",1)))

#plot group means heatmap
jpeg("DMR_Taov0.05_d135mean_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_d135DMR_mean_m,margins = c(10,20), cexRow = 1.2, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)),distfun = function(x) as.dist(1-cor(t(x))), hclustfun = function(x) hclust(x, method="average"), na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_d135DMR_mean_m),rowsep=1:nrow(aov_0.05pH_d135DMR_mean_m))
dev.off()

##### DAY 145  ######
#create matrix for d145 samples
aov_0.05pH_d145DMR_m <- as.matrix(d145DMRs[,7:30])
rownames(aov_0.05pH_d145DMR_m) <- d145DMRs$ID

aov_0.05pH_d145DMR_m <- aov_0.05pH_d145DMR_m[which(rownames(aov_0.05pH_d145DMR_m) %in% pull(d145DMR_1way_aov_pH_modelsumm[which(d145DMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),]
#remove extra text from column names
colnames(aov_0.05pH_d145DMR_m) <- gsub("methylation_level_","",colnames(aov_0.05pH_d145DMR_m))


#order columns by pH treatment (amb-amb, amb-low,low-amb, low-low, superlow-amb, superlow-low)
aov_0.05pH_d145DMR_m <- aov_0.05pH_d145DMR_m[,c(3,4,5,6,13,14,21,22,1,2,9,10,15,16,23,24,7,8,11,12,17,18,19,20)]

#define color bar corresponding to treatment
ColSideColors <- cbind(pH = c(rep("cyan3",4),rep("green3",4),rep("gold2",4),rep("greenyellow",4),rep("magenta3",4),rep("lightslateblue",4)))

jpeg("DMR_Taov0.05_d145_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_d145DMR_m,margins = c(10,20), cexRow = 1.2, cexCol = 1, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)), ColSideColors = ColSideColors,na.color = "black", density.info = "none", trace = "none", scale = "row", sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_d145DMR_m),rowsep=1:nrow(aov_0.05pH_d145DMR_m))
dev.off()

###Visualize group means
#calculate group means

Meand145pHambamb <- rowMeans(aov_0.05pH_d145DMR_m[,grep("181|182|184|185", colnames(aov_0.05pH_d145DMR_m))], na.rm = TRUE)

Meand145pHamblow <- rowMeans(aov_0.05pH_d145DMR_m[,grep("205|206|226|227", colnames(aov_0.05pH_d145DMR_m))], na.rm = TRUE)

Meand145pHlowamb <- rowMeans(aov_0.05pH_d145DMR_m[,grep("175|176|193|194", colnames(aov_0.05pH_d145DMR_m))], na.rm = TRUE)

Meand145pHlowlow <- rowMeans(aov_0.05pH_d145DMR_m[,grep("208|209|229|230", colnames(aov_0.05pH_d145DMR_m))], na.rm = TRUE)

Meand145pHSlowamb <- rowMeans(aov_0.05pH_d145DMR_m[,grep("187|188|199|200", colnames(aov_0.05pH_d145DMR_m))], na.rm = TRUE)

Meand145pHSlowlow <- rowMeans(aov_0.05pH_d145DMR_m[,grep("214|215|220|221", colnames(aov_0.05pH_d145DMR_m))], na.rm = TRUE)

#bind all group means together
aov_0.05pH_d145DMR_mean_m <- as.matrix(data.frame(cbind(Meand145pHambamb,Meand145pHamblow,Meand145pHlowamb,Meand145pHlowlow,Meand145pHSlowamb,Meand145pHSlowlow)))
#define color bar for group means
ColSideColors <- cbind(pH = c(rep("cyan3",1),rep("green3",1),rep("gold2",1),rep("greenyellow",1),rep("magenta3",1),rep("lightslateblue",1)))
#plot group means heatmap
jpeg("DMR_Taov0.05_d145mean_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_d145DMR_mean_m,margins = c(10,20), cexRow = 1.2, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)), distfun = function(x) as.dist(1-cor(t(x))), hclustfun = function(x) hclust(x, method="average"),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_d145DMR_mean_m),rowsep=1:nrow(aov_0.05pH_d145DMR_mean_m))
dev.off()

##### AMBIENT SAMPLES ######

#create matrix for amb samples
aov_0.05pH_ambDMR_m <- as.matrix(ambDMRs[,7:22])
rownames(aov_0.05pH_ambDMR_m) <- ambDMRs$ID

aov_0.05pH_ambDMR_m <- aov_0.05pH_ambDMR_m[which(rownames(aov_0.05pH_ambDMR_m) %in% pull(ambDMR_1way_aov_pH_modelsumm[which(ambDMR_1way_aov_pH_modelsumm$p.value < 0.05),],ID)),]
#remove extra text from column names
colnames(aov_0.05pH_ambDMR_m) <- gsub("methylation_level_","",colnames(aov_0.05pH_ambDMR_m))


#columns are already ordered by time (day0, day10,day135,day145)
#define color bar corresponding to treatment
ColSideColors <- cbind(day = c(rep("gray82",4),rep("gray53",4),rep("gray30",4),rep("gray8",4)))

jpeg("DMR_Taov0.05_amb_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_ambDMR_m,margins = c(10,20), cexRow = 1.2, cexCol = 1, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)), ColSideColors = ColSideColors,na.color = "black", density.info = "none", trace = "none", scale = "row", sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_ambDMR_m),rowsep=1:nrow(aov_0.05pH_ambDMR_m))
dev.off()

###Visualize group means
#calculate group means
Meanambd0 <- rowMeans(aov_0.05pH_ambDMR_m[,grep("41|42|43|44", colnames(aov_0.05pH_ambDMR_m))], na.rm = TRUE)

Meanambd10 <- rowMeans(aov_0.05pH_ambDMR_m[,grep("119|120|135|136", colnames(aov_0.05pH_ambDMR_m))], na.rm = TRUE)


Meanambd135 <- rowMeans(aov_0.05pH_ambDMR_m[,grep("151|152|153|154", colnames(aov_0.05pH_ambDMR_m))], na.rm = TRUE)

Meanambd145 <- rowMeans(aov_0.05pH_ambDMR_m[,grep("181|182|184|185", colnames(aov_0.05pH_ambDMR_m))], na.rm = TRUE)

#bind all group means together
aov_0.05pH_ambDMR_mean_m <- as.matrix(data.frame(cbind(Meanambd0,Meanambd10, Meanambd135,Meanambd145)))
#define color bar for group means
ColSideColors <- cbind(day = c(rep("gray82",1),rep("gray53",1),rep("gray30",1),rep("gray8",1)))

#plot group means heatmap
jpeg("DMR_Taov0.05_ambmean_heatmap.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_ambDMR_mean_m,margins = c(10,20), cexRow = 1.2, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = rev(RColorBrewer::brewer.pal(name = "RdBu", n = 10)), distfun = function(x) as.dist(1-cor(t(x))), hclustfun = function(x) hclust(x, method="average"),na.color = "black", density.info = "none", trace = "none", scale = "row",sepwidth=c(0.025,0.025), sepcolor="darkgray",colsep=1:ncol(aov_0.05pH_ambDMR_mean_m),rowsep=1:nrow(aov_0.05pH_ambDMR_mean_m))
dev.off()
```

HEATMAP OF GROUP MEANS (like Hollie's)
```{r}
#plot for ambient group mean comparison
colnames(aov_0.05pH_ambDMR_mean_m) <- c("Day 0", "Day 10", "Day 135", "Day 145")

jpeg("amb_MCmax25DMR_Taov0.05_GROUPmean_heatmap2.jpg", width = 700, height = 1000)
heatmap.2(aov_0.05pH_ambDMR_mean_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)), na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov_0.05pH_ambDMR_mean_m),rowsep=1:nrow(aov_0.05pH_ambDMR_mean_m))
dev.off()

#plot for day10 group mean comparison
colnames(aov_0.05pH_d10DMR_mean_m) <- c("pH 8.0", "pH 7.4", "pH 7.0")

jpeg("d10_MCmax25DMR_Taov0.05_GROUPmean_heatmap2.jpg", width = 600, height = 1000)
heatmap.2(aov_0.05pH_d10DMR_mean_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)), na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov_0.05pH_d10DMR_mean_m),rowsep=1:nrow(aov_0.05pH_ambDMR_mean_m))
dev.off()

#plot for day135 group mean comparison
colnames(aov_0.05pH_d135DMR_mean_m) <- c("pH 8.0", "pH 7.4", "pH 7.0")

jpeg("d135_MCmax25DMR_Taov0.05_GROUPmean_heatmap2.jpg", width = 600, height = 1000)
heatmap.2(aov_0.05pH_d135DMR_mean_m,margins = c(10,20), cexCol = 2, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)), na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov_0.05pH_d135DMR_mean_m),rowsep=1:nrow(aov_0.05pH_ambDMR_mean_m))
dev.off()

#plot for day145 group mean comparison
colnames(aov_0.05pH_d145DMR_mean_m) <- c("pH 8.0 -> pH 8.0", "pH 8.0 -> pH 7.4","pH 7.4 -> pH 8.0","pH 7.4 -> pH 7.4", "pH 7.0 -> pH 8.0", "pH 7.0 -> pH 7.4")

jpeg("d145_MCmax25DMR_Taov0.05_GROUPmean_heatmap2.jpg", width = 800, height = 1000)
heatmap.2(aov_0.05pH_d145DMR_mean_m,margins = c(12,15), cexCol = 1.5, distfun = function(x) as.dist(1 - cor(t(x), use = "pa")), hclustfun = function(x) hclust(x,method = 'average'),Colv=NA, col= rev(colorRampPalette(brewer.pal(10, "RdYlBu"))(256)), na.color = "black", density.info = "none", trace = "none", scale = "row", labRow = FALSE,sepwidth=c(0.01,0.01),sepcolor="white",colsep=1:ncol(aov_0.05pH_d145DMR_mean_m),rowsep=1:nrow(aov_0.05pH_ambDMR_mean_m))
dev.off()
```


Write out bedfiles
```{r}
#write out d10 DMR bed file
aov_0.05pH_d10DMR_bed <- data.frame(aov_0.05pH_d10DMR_m)
aov_0.05pH_d10DMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_d10DMR_m))
aov_0.05pH_d10DMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_d10DMR_m))
aov_0.05pH_d10DMR_bed$start <- gsub("-.*","", aov_0.05pH_d10DMR_bed$start)
aov_0.05pH_d10DMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_d10DMR_m))
aov_0.05pH_d10DMR_bed$comparison <- "d10"
aov_0.05pH_d10DMR_bed <- aov_0.05pH_d10DMR_bed[,c((ncol(aov_0.05pH_d10DMR_bed)-3):ncol(aov_0.05pH_d10DMR_bed)) ]
write.table(aov_0.05pH_d10DMR_bed,"aov_0.05pH_d10DMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")

#write out d135 DMR bed file
aov_0.05pH_d135DMR_bed <- data.frame(aov_0.05pH_d135DMR_m)
aov_0.05pH_d135DMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_d135DMR_m))
aov_0.05pH_d135DMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_d135DMR_m))
aov_0.05pH_d135DMR_bed$start <- gsub("-.*","", aov_0.05pH_d135DMR_bed$start)
aov_0.05pH_d135DMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_d135DMR_m))
aov_0.05pH_d135DMR_bed$comparison <- "d135"
aov_0.05pH_d135DMR_bed <- aov_0.05pH_d135DMR_bed[,c((ncol(aov_0.05pH_d135DMR_bed)-3):ncol(aov_0.05pH_d135DMR_bed)) ]
write.table(aov_0.05pH_d135DMR_bed,"aov_0.05pH_d135DMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")

#write out d145 DMR bed file
aov_0.05pH_d145DMR_bed <- data.frame(aov_0.05pH_d145DMR_m)
aov_0.05pH_d145DMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_d145DMR_m))
aov_0.05pH_d145DMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_d145DMR_m))
aov_0.05pH_d145DMR_bed$start <- gsub("-.*","", aov_0.05pH_d145DMR_bed$start)
aov_0.05pH_d145DMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_d145DMR_m))
aov_0.05pH_d145DMR_bed$comparison <- "d145"
aov_0.05pH_d145DMR_bed <- aov_0.05pH_d145DMR_bed[,c((ncol(aov_0.05pH_d145DMR_bed)-3):ncol(aov_0.05pH_d145DMR_bed)) ]
write.table(aov_0.05pH_d145DMR_bed,"aov_0.05pH_d145DMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")


#write out amb DMR bed file
aov_0.05pH_ambDMR_bed <- data.frame(aov_0.05pH_ambDMR_m)
aov_0.05pH_ambDMR_bed$Scaffold <- gsub("\\:.*","", rownames(aov_0.05pH_ambDMR_m))
aov_0.05pH_ambDMR_bed$start <- gsub(".*\\:","", rownames(aov_0.05pH_ambDMR_m))
aov_0.05pH_ambDMR_bed$start <- gsub("-.*","", aov_0.05pH_ambDMR_bed$start)
aov_0.05pH_ambDMR_bed$end <- gsub(".*-","",  rownames(aov_0.05pH_ambDMR_m))
aov_0.05pH_ambDMR_bed$comparison <- "amb"
aov_0.05pH_ambDMR_bed <- aov_0.05pH_ambDMR_bed[,c((ncol(aov_0.05pH_ambDMR_bed)-3):ncol(aov_0.05pH_ambDMR_bed)) ]
write.table(aov_0.05pH_ambDMR_bed,"aov_0.05pH_ambDMR.bed", row.names = FALSE,col.names = FALSE, quote = FALSE,sep = "\t")
```

