---
title: "Untitled"
author: "Shelly Trigg"
date: "3/12/2020"
output: html_document
---


load libraries
```{r}
library(tidyr)
library(dplyr)
library(broom)
library(ggplot2)
library(arm)


```

read in data
```{r}
#table with summed mCs and tCs for DMRs
amb_DMR_mC_tC_counts_SUM <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20200311/extractDMS/amb_AllTimes_DMR250bp_MCmax25_cov5x_DMR_SUMMED_mC_tC.tsv", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

#table with mC and tC counts for each DMR
amb_DMR_mC_tC_counts <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20200311/extractDMS/amb_AllTimes_DMR250bp_MCmax25_cov5x_DMR_and_DMS_mC_tC.tsv", header = FALSE, sep = "\t", stringsAsFactors = FALSE)


#header info
amb_DMR_mC_tC_counts_header <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20200311/extractDMS/amb_AllTimes_DMR250bp_MCmax25_cov5x_rms_results.HEADER.tsv", header = FALSE, sep = "\t", stringsAsFactors = FALSE)

#meta data
meta_data <- read.csv("~/Documents/GitHub/Shelly_Pgenerosa/analyses/20191222/Sample.Info.csv", stringsAsFactors = FALSE)
```

Remove empty column in SUM data frame
```{r}
amb_DMR_mC_tC_counts_SUM[,37] <- NULL
```

Remove duplicate columns in count data frame
```{r}
amb_DMR_mC_tC_counts <- amb_DMR_mC_tC_counts[,-c(11:15)]
```



Add colnames
```{r}
colnames(amb_DMR_mC_tC_counts_SUM) <- c("chr","start", "end", "numDMS", amb_DMR_mC_tC_counts_header[1,6:37])


colnames(amb_DMR_mC_tC_counts) <- c("DMRchr","DMRstart", "DMRend", "numDMS","DMSchr","DMSstart", "DMSend","context","rmsPval","BedclosestDist", amb_DMR_mC_tC_counts_header[1,6:37])
```


Rearrange table
```{r}
#add sample name to column and mC to column
amb_DMR_mC_tC_counts_SUM_mC_stacked <- gather(amb_DMR_mC_tC_counts_SUM[,1:20], "Sample.ID", "mC",5:20)
#remove mc from sample column
amb_DMR_mC_tC_counts_SUM_mC_stacked$Sample.ID <- gsub("mc_","",amb_DMR_mC_tC_counts_SUM_mC_stacked$Sample.ID )

#add sample name to column and tC to column
amb_DMR_mC_tC_counts_SUM_tC_stacked <- gather(amb_DMR_mC_tC_counts_SUM[,-c(5:20)], "Sample.ID", "tC",5:20)

amb_DMR_mC_tC_counts_SUM_tC_stacked$Sample.ID <- gsub("h_","",amb_DMR_mC_tC_counts_SUM_tC_stacked$Sample.ID )

#merge data
amb_DMR_mC_tC_counts_SUM_stacked <- merge(amb_DMR_mC_tC_counts_SUM_mC_stacked,amb_DMR_mC_tC_counts_SUM_tC_stacked, by = c("chr", "start", "end", "numDMS", "Sample.ID"))

#convert sample ID column hyphen to underscore
amb_DMR_mC_tC_counts_SUM_stacked$Sample.ID <- gsub("-","_", amb_DMR_mC_tC_counts_SUM_stacked$Sample.ID)


######
#Do the same for count data
######
#add sample name to column and mC to column
amb_DMR_mC_tC_counts_mC_stacked <- gather(amb_DMR_mC_tC_counts[,1:26], "Sample.ID", "mC",11:26)
#remove mc from sample column
amb_DMR_mC_tC_counts_mC_stacked$Sample.ID <- gsub("mc_","",amb_DMR_mC_tC_counts_mC_stacked$Sample.ID )

#add sample name to column and tC to column
amb_DMR_mC_tC_counts_tC_stacked <- gather(amb_DMR_mC_tC_counts[,-c(11:26)], "Sample.ID", "tC",11:26)

amb_DMR_mC_tC_counts_tC_stacked$Sample.ID <- gsub("h_","",amb_DMR_mC_tC_counts_tC_stacked$Sample.ID )

#merge data
amb_DMR_mC_tC_counts_stacked <- merge(amb_DMR_mC_tC_counts_mC_stacked,amb_DMR_mC_tC_counts_tC_stacked, by = c("DMRchr","DMRstart", "DMRend", "numDMS","DMSchr","DMSstart", "DMSend","context","rmsPval","BedclosestDist", "Sample.ID"))

#convert sample ID column hyphen to underscore
amb_DMR_mC_tC_counts_stacked$Sample.ID <- gsub("-","_", amb_DMR_mC_tC_counts_stacked$Sample.ID)


```

Convert mC and tC values between 0 and 1 to NA; these values mean there was no coverage for any DMS within the DMR
```{r}
#convert -1 in counts data to NA
amb_DMR_mC_tC_counts_stacked[amb_DMR_mC_tC_counts_stacked == -1] <- NA


#convert zeros to something unique
amb_DMR_mC_tC_counts_SUM_stacked[amb_DMR_mC_tC_counts_SUM_stacked == 0] <- 50000000

#round down to get integers and convert the integers without coverage (where counts are < 1 and > 0) to zero
amb_DMR_mC_tC_counts_SUM_stacked$mC <- floor(amb_DMR_mC_tC_counts_SUM_stacked$mC)
amb_DMR_mC_tC_counts_SUM_stacked$tC <- floor(amb_DMR_mC_tC_counts_SUM_stacked$tC)
#convert zeros to NA since these had no coverage
amb_DMR_mC_tC_counts_SUM_stacked[amb_DMR_mC_tC_counts_SUM_stacked == 0] <- NA
#convert original zeros back to zero
amb_DMR_mC_tC_counts_SUM_stacked[amb_DMR_mC_tC_counts_SUM_stacked == 50000000] <- 0

#NAs should only exist when there was no coverage of any DMS within DMRs; both mC and tC were initially between 0 and 1. There are now instances where mC has NA and tC does not. This is because for the DMR, some DMS had coverage and some did not AND for the DMS that had coverage none were methylated. The code below fixes these NAs.
for (i in 1:nrow(amb_DMR_mC_tC_counts_SUM_stacked)){
  if(is.na(amb_DMR_mC_tC_counts_SUM_stacked$mC[i])& !(is.na(amb_DMR_mC_tC_counts_SUM_stacked$tC[i]))){
    amb_DMR_mC_tC_counts_SUM_stacked$mC[i] <- 0
  }
}
```

Create a column with number of non-methylated Cs
```{r}
amb_DMR_mC_tC_counts_SUM_stacked$nmC <- (amb_DMR_mC_tC_counts_SUM_stacked$tC - amb_DMR_mC_tC_counts_SUM_stacked$mC)


amb_DMR_mC_tC_counts_stacked$nmC <- (amb_DMR_mC_tC_counts_stacked$tC - amb_DMR_mC_tC_counts_stacked$mC)
```

Merge counts data with meta data
```{r}
amb_DMR_mC_tC_counts_SUM_stacked <- merge(meta_data, amb_DMR_mC_tC_counts_SUM_stacked, by = "Sample.ID")

amb_DMR_mC_tC_counts_stacked <- merge(meta_data, amb_DMR_mC_tC_counts_stacked, by = "Sample.ID")
```


Create a column for DMR ID
```{r}
amb_DMR_mC_tC_counts_SUM_stacked$DMR.ID <- paste(amb_DMR_mC_tC_counts_SUM_stacked$chr,amb_DMR_mC_tC_counts_SUM_stacked$start, amb_DMR_mC_tC_counts_SUM_stacked$end, sep = "_")


amb_DMR_mC_tC_counts_stacked$DMR.ID <- paste(amb_DMR_mC_tC_counts_stacked$DMRchr,amb_DMR_mC_tC_counts_stacked$DMRstart, amb_DMR_mC_tC_counts_stacked$DMRend, sep = "_")
```

Create a column for unique DMS position in counts data
```{r}
amb_DMR_mC_tC_counts_stacked$DMS.position <- paste(amb_DMR_mC_tC_counts_stacked$DMSchr,amb_DMR_mC_tC_counts_stacked$DMSstart, sep = "_")
```

Run the glm

```{r}
ambDMR_glm <- amb_DMR_mC_tC_counts_SUM_stacked %>% group_by(DMR.ID) %>%
do(meth_glms = glm(matrix(c(mC,nmC), ncol =2) ~ TimePoint, data =  . , family= binomial))

ambDMR_glm_pvals <- amb_DMR_mC_tC_counts_SUM_stacked %>% group_by(DMR.ID) %>%
do(meth_glm_pvals = anova(glm(matrix(c(mC,nmC), ncol =2) ~ TimePoint, data =  . , family= binomial), test = "Chisq"))

```

Summarize glm data
```{r}
ambDMR_glm_pvals_modelsumm <- tidy(ambDMR_glm_pvals,meth_glm_pvals)
ambDMR_glm_pvals_modelsumm <- ambDMR_glm_pvals_modelsumm[which(ambDMR_glm_pvals_modelsumm$term != "NULL"),]

```

FDR correct glm data
```{r}
ambDMR_glm_pvals_modelsumm$p.adj <- p.adjust(ambDMR_glm_pvals_modelsumm$p.value, method = "BH")


```


What is the distribution of the count data?
```{r}
ggplot(amb_DMR_mC_tC_counts_SUM_stacked[which(amb_DMR_mC_tC_counts_SUM_stacked$mC <1000),]) + geom_histogram(aes(mC, group = TimePoint, color = TimePoint,fill = TimePoint), bins = 10, position = "identity", alpha = 0.5) + theme_bw()

#plot the data excluding DMRs with coverage over 1000 reads
ggplot(amb_DMR_mC_tC_counts_SUM_stacked[which(amb_DMR_mC_tC_counts_SUM_stacked$mC <1000),]) + geom_histogram(aes(mC, group = TimePoint, color = TimePoint,fill = TimePoint), bins = 10, position = "identity", alpha = 0.5) + theme_bw()

```

The data are definately poisson distributed suggesting a poisson glm should be run
```{r}
ambDMR_pois_glm <- amb_DMR_mC_tC_counts_SUM_stacked %>% group_by(DMR.ID) %>%
do(meth_pois_glms = glm(mC ~ TimePoint, data =  . , family= poisson))

ambDMR_pois_glm_pvals <- amb_DMR_mC_tC_counts_SUM_stacked %>% group_by(DMR.ID) %>%
do(meth_pois_glm_pvals = anova(glm(mC ~ TimePoint, data =  . , family= poisson), test = "Chisq"))


ambDMR_pois_glm_pvals_modelsumm <- tidy(ambDMR_pois_glm_pvals,meth_pois_glm_pvals)
ambDMR_pois_glm_pvals_modelsumm <- ambDMR_pois_glm_pvals_modelsumm[which(ambDMR_pois_glm_pvals_modelsumm$term != "NULL"),]


ambDMR_pois_glm_pvals_modelsumm$p.adj <- p.adjust(ambDMR_pois_glm_pvals_modelsumm$p.value, method = "BH")

```


Run the glm including position 

```{r}
ambDMR_glm <- amb_DMR_mC_tC_counts_stacked %>% group_by(DMR.ID) %>%
do(meth_glms = glm(matrix(c(mC,nmC), ncol =2) ~ TimePoint * DMS.position, data =  . , family= binomial))

ambDMR_glm_pvals <- amb_DMR_mC_tC_counts_stacked %>% group_by(DMR.ID) %>%
do(meth_glm_pvals = anova(glm(matrix(c(mC,nmC), ncol =2) ~ TimePoint * DMS.position, data =  . , family= binomial), test = "Chisq"))

```

Summarize glm including position data
```{r}
ambDMR_glm_pvals_modelsumm <- tidy(ambDMR_glm_pvals,meth_glm_pvals)
ambDMR_glm_pvals_modelsumm <- ambDMR_glm_pvals_modelsumm[which(ambDMR_glm_pvals_modelsumm$term != "NULL"),]

```

FDR correct glm data
```{r}
ambDMR_glm_pvals_modelsumm$p.adj <- p.adjust(ambDMR_glm_pvals_modelsumm$p.value, method = "BH")
```

Still getting pvalues that don't make sense (everything is significant)
I'm going to try bayesglm as suggested here https://discuss.analyticsvidhya.com/t/glm-fit-fitted-probabilities-numerically-0-or-1-occurred-warning-message-when-i-run-logistic-regression/10390/3
```{r}
ambDMR_bayesglm <- amb_DMR_mC_tC_counts_stacked %>% group_by(DMR.ID) %>%
do(meth_bayesglms = bayesglm(matrix(c(mC,nmC), ncol =2) ~ TimePoint * DMS.position, data =  . , family= "binomial"))

ambDMR_bayesglm_pvals <- amb_DMR_mC_tC_counts_stacked %>% group_by(DMR.ID) %>%
do(meth_bayesglm_pvals = anova(bayesglm(matrix(c(mC,nmC), ncol =2) ~ TimePoint * DMS.position, data =  . , family= "binomial"), test = "Chisq"))
```

Summarize bayesglm data
```{r}
ambDMR_bayesglm_pvals_modelsumm <- tidy(ambDMR_bayesglm_pvals,meth_bayesglm_pvals)
ambDMR_bayesglm_pvals_modelsumm <- ambDMR_bayesglm_pvals_modelsumm[which(ambDMR_bayesglm_pvals_modelsumm$term != "NULL"),]

```

FDR correct glm data
```{r}
ambDMR_bayesglm_pvals_modelsumm$p.adj <- p.adjust(ambDMR_bayesglm_pvals_modelsumm$p.value, method = "BH")

```
