---
title: "DMR Heatmaps"
author: "Shelly Trigg"
date: "9/30/2019"
output: rmarkdown::github_document
---

This script was run with Gannet mounted


load libraries
```{r}
library(gplots)

```

read in data
```{r}
sep26_DMRs_allAmb <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20190926/amb_AllTimes_DMR250bp_MCmax25_cov5x_clst3_rms_results_collapsed.tsv", header = TRUE, sep = "\t")
sep26_DMRs_day10 <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20190926/day10_AllpH_DMR250bp_MCmax25_cov5x_clst3_rms_results_collapsed.tsv", header = TRUE, sep = "\t")
sep26_DMRs_day135 <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20190926/day135_AllpH_DMR250bp_MCmax25_cov5x_clst3_rms_results_collapsed.tsv", header = TRUE, sep = "\t")
sep26_DMRs_day145 <- read.table("/Volumes/web/metacarcinus/Pgenerosa/analyses/20190926/day145_AllpH_DMR250bp_MCmax25_cov5x_clst3_rms_results_collapsed.tsv", header = TRUE, sep = "\t")

```

Make a unique ID column in each data frame
```{r}
#for all ambient sample comparison
sep26_DMRs_allAmb$ID <- paste(sep26_DMRs_allAmb$chr,":",sep26_DMRs_allAmb$start,"-",sep26_DMRs_allAmb$end, sep = "")
sep26_DMRs_allAmb$ID <- gsub("__.*__.*:",":",sep26_DMRs_allAmb$ID)

#for day 10 sample comparison

sep26_DMRs_day10$ID <- paste(sep26_DMRs_day10$chr,":",sep26_DMRs_day10$start,"-",sep26_DMRs_day10$end, sep = "")
sep26_DMRs_day10$ID <- gsub("__.*__.*:",":",sep26_DMRs_day10$ID)

#for day 135 sample comparison

sep26_DMRs_day135$ID <- paste(sep26_DMRs_day135$chr,":",sep26_DMRs_day135$start,"-",sep26_DMRs_day135$end, sep = "")
sep26_DMRs_day135$ID <- gsub("__.*__.*:",":",sep26_DMRs_day135$ID)

#for day 145 sample comparison

sep26_DMRs_day145$ID <- paste(sep26_DMRs_day145$chr,":",sep26_DMRs_day145$start,"-",sep26_DMRs_day145$end, sep = "")
sep26_DMRs_day145$ID <- gsub("__.*__.*:",":",sep26_DMRs_day145$ID)


```

create matrix for all ambient samples
```{r}
allAmb_m <- as.matrix(sep26_DMRs_allAmb[,7:22])
rownames(allAmb_m) <- sep26_DMRs_allAmb$ID
```

plot heatmap of all ambient sample comparison
```{r, allAmb_DMR_heatmap, fig.height= 8, fig.width= 11}
ColSideColors <- cbind(day = c(rep("#969696",4),rep("#737373",4),rep("#525252",4),rep("#252525",4)))
heatmap.2(allAmb_m,cexRow = 0.3, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = bluered, na.color = "black", density.info = "none", trace = "none", scale = "row")
```

create matrix for day10 samples
```{r}
#subset out the data and order it
day10_m <- as.matrix(sep26_DMRs_day10[,c(11:14,7:8,15:16,9:10,17:18)])
rownames(day10_m) <- sep26_DMRs_day10$ID
```

plot heatmap of day10 sample comparison
```{r, day10_DMR_heatmap}
ColSideColors <- cbind(pH = c(rep("cyan",4),rep("plum2",4),rep("magenta",4)))
heatmap.2(day10_m,cexRow = 0.3, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = bluered, na.color = "black", density.info = "none", trace = "none", scale = "row")
```

create matrix for day135 samples
```{r}
#subset out the data and order it
day135_m <- as.matrix(sep26_DMRs_day135[,c(7:18)])
rownames(day135_m) <- sep26_DMRs_day135$ID
```

plot heatmap of day135 sample comparison
```{r, day135_DMR_heatmap}
ColSideColors <- cbind(pH = c(rep("cyan",4),rep("plum2",4),rep("magenta",4)))
heatmap.2(day135_m,cexRow = 0.3, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = bluered, na.color = "black", density.info = "none", trace = "none", scale = "row")
```


create matrix for day145 samples
```{r}
#subset out the data and order it
day145_m <- as.matrix(sep26_DMRs_day145[,c(9:12,7:8,15:16,13:14,17:18,19:20,27:28,21:22,29:30,23:26)])
rownames(day145_m) <- sep26_DMRs_day145$ID
```

plot heatmap of day145 sample comparison
```{r, day145_DMR_heatmap}
ColSideColors <- cbind(pH = c(rep("cyan",4),rep("green1",4),rep("green3",4),rep("mediumpurple1",4),rep("plum2",4),rep("magenta",4)))
heatmap.2(day145_m,cexRow = 0.3, cexCol = 1,ColSideColors = ColSideColors, Colv=NA, col = bluered, na.color = "black", density.info = "none", trace = "none", scale = "row")

