---
title: "Get_DMRs_for_SR_v074_alignments"
output: html_document
---

Load libraries
```{r}
library(methylKit)
library(ggplot2)

```

create file list
```{r}
file.list=c(dir("~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/dedup_bams/",pattern = ".sorted.bam"))
file.list <- paste("~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/dedup_bams/", file.list, sep = "")
file.list <- as.list(file.list)
```


read in data
```{r}
treatment <- read.csv("~/GitHub/Shelly_Pgenerosa/data/Treatment_info_CORRECTED.csv", header = TRUE)

```

make a list of samples that are only from 145 days Ambient-low or Super.Low-low
```{r}
select_treats <- treatment[which((treatment$Treatment == "Super.Low-Low" & treatment$TimePoint == "Day145")|(treatment$Treatment == "Ambient-Low" & treatment$TimePoint == "Day145")),]

li <- as.list(gsub("~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/dedup_bams/EPI-","",file.list))
li <- as.list(gsub("_.*","",li))

```

```{r}
myobj <- processBismarkAln(location = file.list, sample.id = li, assembly = "v3", read.context="CpG", mincov=3, treatment = select_treats$Treatment)

save(myobj, file = "~/GitHub/Shelly_Pgenerosa/data/20190816_myobj")


#load("~/GitHub/Shelly_Pgenerosa/data/20190816_myobj")
```

calculate region coverage 
```{r}
#step below took ~24 hours

tiles <- tileMethylCounts(myobj,win.size=1000,step.size=1000,mc.cores = 3)
save(tiles, file = "~/GitHub/Shelly_Pgenerosa/data/20190816_tiles")

#load("~/GitHub/Shelly_Pgenerosa/data/20190816_tiles")

```

individual plots of % methylation per region  

```{r, echo = FALSE}
for(i in 1:length(tiles)){
  getMethylationStats(tiles[[i]],plot = TRUE, both.strands = FALSE)
}
```

individual plots of read coverage per region  

```{r, echo = FALSE}
for(i in 1:length(tiles)){
  getCoverageStats(tiles[[i]],plot = TRUE, both.strands = FALSE)
}
```


find regions covered by all samples (This finds regions that are covered by two or more samples I believe, because the covereage has NAs for some samples.)
```{r}
mmeth <- unite(tiles, min.per.group = 1L)
```
sample clustering by differences in correlation of methylation patterns between samples
```{r, echo = FALSE}
clusterSamples(mmeth, dist = "correlation", method = "complete", plot = TRUE)
```

PCA of samples based on methylation patterns
```{r, echo = FALSE}
jpeg(filename = "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/PCA.jpeg", height = 1000, width = 1000) #Save file with designated name
PCASamples(mmeth) #Run the PCA analysis and plot variances against PC number in a screeplot
dev.off()

```


confirm treatment and control groups are correctly assigned
```{r}
#view treatment assignments in the unite output
mmeth
#view the treatment assignments in the initial table
select_treats$Treatment

#confirmed treatment and contrl groups are correctly assigned
```


make a file with diff meth
```{r}

myDiff.loci <- calculateDiffMeth(mmeth, mc.cores = 3)

myDiff50p.l=getMethylDiff(myDiff.loci,difference=50,qvalue=0.01)
write.table(getData(myDiff50p.l), "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/SR_v074_ALvSL-L_hyperANDhypoDMR.tsv", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)



# get hyper-methylated  
hyper.l=getMethylDiff(myDiff.loci,difference=50,qvalue=0.01,type="hyper")
write.table(getData(hyper.l), "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/SR_v074_ALvSL-L_hyperDMR.tsv", sep = "\t", quote = FALSE, row.names = FALSE)

# get hypo-methylated
hypo.l=getMethylDiff(myDiff.loci,difference=50,qvalue=0.01,type="hypo")
write.table(getData(hypo.l), "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/SR_v074_ALvSL-L_hypoDMR.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
```

visualize
```{r}
jpeg(filename = "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/diffmethbychrom.jpeg", height = 1000, width = 1500) #Save file with designated name
diffMethPerChr(myDiff.loci,qvalue.cutoff=0.01, meth.cutoff=50)
dev.off()

diffMethPerChr <-diffMethPerChr(myDiff.loci,qvalue.cutoff=0.01, meth.cutoff=50)

```