---
title: "Get_DMRs_for_SR_v074_alignments_unite3LdestrandT"
output: html_document
---

#This script is run on Emu's RStudio web interface

Load libraries
```{r}
library(methylKit)
library(ggplot2)

```


```{r}
load("~/GitHub/Shelly_Pgenerosa/data/20190816_myobj")
```

calculate region coverage 
```{r}
load("~/GitHub/Shelly_Pgenerosa/data/20190816_tiles")
```


find regions covered by all samples (This finds regions that are covered by 3 of 4 samples within one treatment and for non-covered C's it assigns NAs). Also create unite object without specifying min.per.group. The min.per.group specification is an integer denoting minimum number of samples per replicate needed to cover a region/base. By default only regions/bases that are covered in all samples are united as methylBase object
```{r}
mmeth <- unite(tiles, destrand = TRUE, min.per.group = 3L, mc.cores = 3)


mmeth_nomin <- unite(tiles, destrand = TRUE, mc.cores = 3)
```


sample clustering by differences in correlation of methylation patterns between samples
```{r, echo = FALSE}
clusterSamples(mmeth, dist = "correlation", method = "complete", plot = TRUE)
```

sample clustering by differences in correlation of methylation patterns between samples
```{r, echo = FALSE}
clusterSamples(mmeth_nomin, dist = "correlation", method = "complete", plot = TRUE)
```


PCA of samples based on methylation patterns
```{r, echo = FALSE}
jpeg(filename = "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/PCA_unite3LdestrandT.jpeg", height = 1000, width = 1000) #Save file with designated name
PCASamples(mmeth) #Run the PCA analysis and plot variances against PC number in a screeplot
dev.off()

```


make a file with diff meth
```{r}

myDiff.loci <- calculateDiffMeth(mmeth, mc.cores = 3)

myDiff.loci.nomin <- calculateDiffMeth(mmeth_nomin, mc.cores = 3)


myDiff50p.l=getMethylDiff(myDiff.loci,difference=50,qvalue=0.01)

myDiff50p.l.nomin=getMethylDiff(myDiff.loci,difference=50,qvalue=0.01)


write.table(getData(myDiff50p.l), "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/SR_v074_ALvSL-L_hyperANDhypoDMR_unite3Ldestrand.tsv", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)

write.table(getData(myDiff50p.l.nomin), "~/GitHub/Shelly_Pgenerosa/analyses/JuviPgen_ALSL2lowd145/Get_DMRs_for_SR_v074_alignments/SR_v074_ALvSL-L_hyperANDhypoDMR_unite.minpgdefault.destrand.tsv", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
```